---
title: "What is Data Science? Lecture 1"
output: html_document
date: "2026-05-19"
---
 
#Data Wrangling:
Import -> Tidy -> Transform, Visualize, Model -> Communicate 

#Data Table
Row - subjects
Column - variable 

#Tibbles
Tibbles are data frames 

#Packages: readr, haven, readxl, tidyverse 

####
This is a repository for data wrangling I, part of Data Science I.

I'm an R Markdown document! 

# Section 1

Here's a **code chunk** that samples from 
a _normal distribution_:

```{r}
samp = rnorm(100)
length(samp)
```

# Section 2

I can take the mean of the sample, too!
The mean is `r mean(samp)`.


### loading in the code:

```{r}
library(tidyverse)
```
Loading in the library because functions used below use tidyverse library and to be more particular, they use the readr package. 

##Reading in the data, path - says where the data lives.
Absolute directions - from the very first folder on the computer and following another folders, get the data. 
Relative path - it is an easier, generalized way. This is starting from the project folder. If stuff is within the same space, then it is easy to locate documents/projects/files, etc.

#Read in the litters dataset:
```{r}
litters_df = read_csv("./data_import_examples/FAS_litters.csv")
```
read_csv is within readr within tidyverse
. means the starting point for reading in the data, so the project folder


Process to clean-up the data names:
- Data originally is not in the "snake_case", but written as normal words, so easiest thing to do is to make all the column names a snake_case. Here, we will update the names for the variables identified. 
- We are going to use clean_names - this function is in the janitor package. We can either access the janitor package by loading it in the libraries chunk, or access it using the janitor:: line.
- Now, litters_df will overwrite the previous litters_df 
```{r}
litters_df = janitor::clean_names(litters_df)
```

##Take a look at the data:
- gives number of rows and columns and prints some frist rows and columns
- everything is a snake_case
```{r}
litters_df
```

#Other things that we can do to look at it:
```{r}
head(litters_df)
tail(litters_df)
```

#Skim function:
- comes from skimr package, we can load it by using :: 
- it will display a summary of things 
- if we have a character variable, it will tell me  how many are missing 
- for numeric variable, it will tell me mean, stdev, percentile, values and histogram
```{r}
skimr::skim(litters_df)
```

#View funciton
- used in the console,  not in the code chunk
- it will open a separate, full dataset window with the dataframe when placed in the console

#in this dataset, GROUP is a character variable - mix of letters and numbers, the pther rest of the row names are called a DOUBLE - DOUBLE is special name for a NUMBER 
- missing values in the dataset will be called an NA 








